From physical principles to relativistic classical Hamiltonian and Lagrangian particle mechanics

### 2015-07-15 2015CompleteDerivation/completeDerivation.tex
Starting to think about a new version which includes quantum

>>>>

\section{About}

The work is organized into:
\begin{description}
  \item[Assumptions] these characterize the physical system we are studying and constitute the premise of our discussion. A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or why not) such an assumption makes sense (in a particular case).
  \item[Definitions] these encode into mathematical language the physical concepts we are studying. A \textbf{justification} follows each definition, which uses physical and mathematical arguments to explain why such a definition and its properties are necessary to describe a particular physical concept.
  \item[Propositions] these are statements that mathematically follow from definitions or other propositions. A mathematical \textbf{proof} follows each proposition. No mathematical breakthrough should be expected as we mostly use well known results from different fields. The goal is to show how those results make sense in light of the physical assumptions we start with.
\end{description}
This should allow you, the reader, to focus on the parts you are most interested in and skim the rest. The philosophically inclined may focus on the assumptions, their rationale and the definitions. The mathematically inclined may focus on the definitions, the propositions and their proofs. The scientist on the assumptions, the definitions and their justifications, and the propositions.

\section{New}

%TODO: it would be nice to have "examples" always refer to the same cases, and that they pertain to different type of physics. E.g. trajectory, gravity, thermodynamics, fluid dynamics, biology, electromagnetism...

We start by fixing a physical system we want to study, that is something we can interact with and perform measurements on (e.g. a cat, a planet, water). We set what particular aspect we want to study (e.g. the trajectory when falling, the motion around a star, its flow in a pipe). We call state a particular configuration in time of the portion under study. We call internal dynamics the remainder of the system.

\begin{assump2}
The physical system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
\end{assump2}

\textbf{Rationale}. While the choice of the object of our study (i.e. the system and the aspect we are interested in) is arbitrary, this assumption is at the heart of being able to write physics laws, to define physical states and to physically study a system. If we are to study a physical system through reliable experimentation and observation, we need deterministic and reversible processes to manipulate its state. We can provide different arguments that point in the same direction.

First, to be able to identify the system, we must first be able to isolate it from the environment. Intuitively, we can distinguish between two chairs because we can move the first to another room and sit on it without somehow having to drag the second as well. More precisely, we can manipulate the state of the first system without affecting the second, and vice-versa. If we extend this idea, to be able to properly call something a system we need to show that it is isolated: if I affect its state, the state of the other systems is not affected and vice-versa. In other words: the system future and past states are only determined by its own state.

Second, the aim of physics is to write laws that can be used to make prediction that can be validated experimentally. If I drop an anvil from a tower, it will accelerate at $9.81 m/s^2$; if I want the anvil to reach the ground at $x m/s$ I have to drop it from $y m$. To the extent that we want to make predictions in time, we need to have a one-to-one correspondence between initial and final states

Third, operationally we must reliably prepare and measure states. That is, we need a process for which the input settings of our preparing device determine the outgoing state of the system; and a process for which the incoming state of the system can be reconstructed by the output of the measuring device. That is, our system must participate in a deterministic and reversible process with the preparing and measuring device. How else could we calibrate our experimental apparatus?

This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and system theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}. One source of confusion, though, may come from statistical process: those are not deterministic! The confusion is that the object of our study stops at the distribution (the ensemble), the part that can be indeed calculated, measured and prepared. Another source of confusion is that what is considered state changes depending on the context: the position of a molecule is part of the state for a single molecule system but not for a volume of gas. And again because in the latter context is not part of the deterministic motion of the system.

So, if it seems we need to require deterministic and reversible motion, why we call this an assumption? Why not call it requirement? Because it's a \emph{simplifying} assumption: it can never be totally achieved. A system can be prepared or measured up to a certain level of precision. Perfect isolation of a system is impossible both practically (e.g. black-body radiation, gravity, ...) and conceptually (e.g. if the system is perfectly isolated, we can't prepare it, we can't measure it, we can't study it: it's not a physical system). If the system is always interacting with the environment, then how can we get deterministic and reversible evolution? We can if the environment and the internal dynamics of the system interact in such a way that they affect and are little affected by the aspect we are studying. For example: the trajectory of a falling cat does not greatly affect and is not greatly affected by its digestive system or external temperature.

\textbf{Rationale}. This makes a clear demarcation of what a classical object is (and conversely what a quantum object is not): it is something that can be fully described up to its infinitesimal constituents. This is clearly a \emph{simplifying} assumption, and it is instructive to understand when and why it cannot be used.

The first problem is methodological. As we saw before, we need access to a deterministic and reversible process to be able to study a system and define a state. For two balls, we can imagine studying pieces of the balls in isolation, and then describe the collision between the two by describing what happens at each piece. The classical assumption holds. For an electron and a photon, we cannot take pieces of the electron or the photon, study them in isolation and then describe how each part moves during Compton scattering. The classical assumption does not hold: we do not have suitable physical processes at our disposal.

The second problem is more conceptual. As we saw before, a system cannot be fully isolated from the environment. The smaller the part we study, the more it will be susceptible to fluctuations from the environment, the less we can justify the deterministic assumption, the less we are in a position to define its state. While the overall position of a ball can be considered fairly unaffected by its surrounding, the position of each small piece is going to be constantly affected by air and light scattering on its surface.

If the scale of the problem
